Stack Safety as a Security Property

### 2020-07-10 main.tex

\paragraph*{Capability machines}
%
%Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines such as CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies, including
the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}\bcp{I thought these policies used
  linear capabilities?}\leo{Only the second one AFAIK. What would you
  change to characterize it more accurately?}.  Capabilities package up pointers
(ordinary data) with base and bounds information (policy data),
and they require rewriting source code (e.g., to use capabilities in place
of ordinary pointers), so separating ordinary machine state from policy
state is delicate; we conjecture it can be done with some effort.

@@

\item From caller to callee, when the machine executes an instruction tagged
  with $\tagCall$ as belonging to the call map. At this point, the current
  {\PCname} depth $n$ is incremented to $\tagPCDepth{n + 1}$; on function entry,
  the locations of the newly allocated frame are marked $\tagStackDepth{n + 1}$.

\item From callee back to caller, when execution reaches the corresponding
  return instruction for the call instruction, as defined by
  $\ret{\mach_c}\ \mach_r$. At this point, the {\PCname} tag is decremented.
  Prior to this point, the locations of the old frame have been restored to
  $\tagNoDepth$.

@@

To enforce this discipline, we define blessed header and return instruction
sequences. These are specific sequences of tagged instructions that are intended
to appear at the beginning and end of each function. This is the header
sequence:
%
\setcounter{pcctr}{1}
\[
  \begin{array}{l|l|l|l}
     & \mathit{Instruction} & \mathit{Tags} & \\
    \hline
    \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
        {Store return address in stack}
    \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
        {Increment stack pointer by frame size}
  \end{array}
\]
%
And the return sequence:
%
\setcounter{pcctr}{1}
\[
  \begin{array}{l|l|l|l}
     & \mathit{Instruction} & \mathit{Tags} & \\
    \hline
    \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
        {Load return address}
    \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
        {Decrement stack pointer by frame size}
    \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
        {Return to caller ($\jalr$)}
  \end{array}
\]

@@

These sequences also enforce the $\SP$ discipline. When stack frames
are of fixed size, it suffices to increment $\SP$ and tag it with a
dedicated tag in the header sequence, which is later checked after
decrementing $\SP$ during the exit sequence to authorize the
return. The well-formedness of the blessed sequences can be checked
statically. Only programs whose sequences are well-formed are
protected by the micropolicy, although programs are still protected
(by fail-stopping) even if the intended sequences are missing. By
tracking the execution of the blessed sequences in the tag state of
the {\PCname} register, the micropolicy ensures that the entry
sequences of calls declared in the call map are executed correctly.
%
During normal execution, the {\PCname} uses the current activation
depth in its tag $\tagPCDepth{n}$ to gate load and store operations,
which are allowed to operate only on the corresponding stack frame,
tagged $\tagStackDepth{n}$.
%\rb{Note that we have also considered and implemented more flexible schemes
%  based on allowing operations above the current activation depth. How do we put
%  these together?}
%\rb{Do we want to discuss variations involving frame pointers or more implicit
%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
%\rb{Where to talk about the role of the compiler?}

>>>>

\paragraph*{Capability machines}
%
%Other policies are implemented in hardware or built on hardware mechanisms.
Capability machines such as CHERI \citep{Woodruff+14} extend conventional
architectures to support efficient and fine-grained control over memory
accesses, which can be used to implement security policies, including
the stack protection policies of
\citet{Skorstengaard+19,Skorstengaard+19b}\bcp{I thought these policies used
  linear capabilities?}\leo{Only the second one AFAIK. What would you
  change to characterize it more accurately?}\rb{StkPointers uses linear
  capabilities, its predecessor users local capabilities (and unlike
  StkPointers could be implemented in CHERI)}.
Capabilities package up pointers
(ordinary data) with base and bounds information (policy data),
and they require rewriting source code (e.g., to use capabilities in place
of ordinary pointers), so separating ordinary machine state from policy
state is delicate; we conjecture it can be done with some effort.

@@

\item From caller to callee, when the machine executes an instruction marked as a call in
  the call map. 
  At this point, the current
  $\tagPCDepth{n}$ tag is incremented to $\tagPCDepth{(n + 1)}$, and the function
  entry sequence initializes  all locations of the new function frame with
  tag $\tagStackDepth{(n + 1)}$.

\item From callee back to caller, when execution is about to reach the return point
  $\ret{c}$
  corresponding to the call at $c$.  The exit sequence retags all locations in the function
  frame with $\tagNoDepth$. At the actual return instruction, the
  $\PCname$ tag is decremented.

@@

This discipline suffices to enforce local stack encapsulation.
To implement it,  we define ``blessed'' instruction sequences
intended to appear at the entry and exit of each function,
which manipulate tags as just described in addition to performing the
usual calling convention tasks of saving/restoring the return address to/from
the stack and adjusting the stack pointer. With the aid of an additional tag on
the $\rsp$ register, these sequences also serve to enforce
well-bracketed control flow.
The micropolicy guarantees atomic execution of these sequences
using a combination of tags on the instructions
and an additional tag on the $\PCname$; we omit the details here.

There remains the question of how to ensure that the sequences are
invoked at the right places. For a program to enjoy the stack safety property
(with respect to a particular call map), all we need require is that 
the entry sequence be initiated at any instruction marked as a call
in the call map. This is achieved by giving these instructions (another) special tag,
and it is easy to check statically that this has been done correctly.
If the code fails to initiate an exit
sequence at a point where the stack safety property expects a return,
the micropolicy will incorrectly behave as if execution
is continuing in the callee, but since the callee never has stronger access
rights than the caller, this is harmless.

%% APT: removed because the tag part is too mysterious and the instruction part
%% is standard. 
%% This is the entry
%% sequence:
%% %
%% \setcounter{pcctr}{1}
%% \[
%%   \begin{array}{l|l|l|l}
%%      & \mathit{Instruction} & \mathit{Tags} & \\
%%     \hline
%%     \row{\sw ~ \rsp ~ \tta ~ 1}{[\rsp+1] \leftarrow \rra}{\tagHa, \tagInstr}
%%         {Store return address in stack}
%%     \row{\addi ~ \rsp ~ \rsp ~ 2}{\rsp \leftarrow \rsp + 2}{\tagHb, \tagInstr}
%%         {Increment stack pointer by frame size}
%%   \end{array}
%% \]
%% %
%% And the exit sequence:
%% %
%% \setcounter{pcctr}{1}
%% \[
%%   \begin{array}{l|l|l|l}
%%      & \mathit{Instruction} & \mathit{Tags} & \\
%%     \hline
%%     \row{\lw ~ \rra ~ \rsp ~ \negate 1}{\rra \leftarrow [\rsp - 1]}{\tagRa, \tagInstr}
%%         {Load return address}
%%     \row{\addi ~ \rsp ~ \rsp ~ \negate 2}{\rsp \leftarrow \rsp - 2}{\tagRb, \tagInstr}
%%         {Decrement stack pointer by frame size}
%%     \row{\jalr ~ \rra ~ \rra ~ 0\apt{wrong}}{\PCname \leftarrow \rra}{\tagRc, \tagInstr}
%%         {Return to caller ($\jalr$)}
%%   \end{array}
%% \]

@@

%% These sequences also enforce the $\SP$ discipline. When stack frames
%% are of fixed size, it suffices to increment $\SP$ and tag it with a
%% dedicated tag in the header sequence, which is later checked after
%% decrementing $\SP$ during the exit sequence to authorize the
%% return. The well-formedness of the blessed sequences can be checked
%% statically.

%% Only programs whose sequences are well-formed are
%% protected by the micropolicy, although programs are still protected
%% (by fail-stopping) even if the intended sequences are missing. By
%% tracking the execution of the blessed sequences in the tag state of
%% the {\PCname} register, the micropolicy ensures that the entry
%% sequences of calls declared in the call map are executed correctly.

Note that in this policy both the entry and exit sequences
must write to each element of
the frame, which can be quite expensive, especially for programs that
allocate large but sparsely populated frames; we return to this point
in \cref{sec:lazy}.
%
%\rb{Note that we have also considered and implemented more flexible schemes
%  based on allowing operations above the current activation depth. How do we put
%  these together?}
%\rb{Do we want to discuss variations involving frame pointers or more implicit
%  return addresses, argument passing on the stack, alloc/dealloc, etc.?}
%\rb{Where to talk about the role of the compiler?}
